MATH 245 Homework 3

### 2024-02-23 PDEs/math245_hw3.Rmd
pde hw3 mostly done

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
##Done: 1, 6
##Not started: 2-5

@@

After setting $y=-s, dy=-ds$ in the integral of $\psi$:
$$u(-x,t)= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(-y)dy}=-u(x,t).$$
Therefore, odd initial conditions imply that the solution of the wave equation is odd.

@@

$$u(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy $$
Which means, when $x=-x$:
$$u(-x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(-x-y)^2/4kt}f(y)\;dy $$
Now use the change of variable $y=-s, dy=-ds$:
$$u(-x,t)= \frac{-1}{\sqrt{4 \pi kt}} \int_{\infty}^{-\infty} e^{-(-x+s)^2/4kt}f(-s)\;ds $$

@@

Now assume $f$ is odd. By definition we have $f(-s)=-f(s)$, so

@@

And therefore our solution is odd when our initial $f(x)$ is an odd function. Similarly, if we assume $f(x)$ is even, by definition we have $f(-s)=f(s)$, so
$$u(-x,t)=\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds = 
\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(s)\;ds = u(x,t), $$
meaning our solution is also even.

@@

@@

## Question 3 (Heat Equation with Robin boundary)
Consider the following problem with a Robin boundary condition:
\begin{equation*}
    \begin{cases}
      u_{t}=k u_{xx} & 0 <x<\infty,0 <t<\infty \\
      u(0,x)=x & t=0,0 <x<\infty \\
      u_x(0,t)-2u(0,t)=0 & x=0
    \end{cases}       
\end{equation*}

@@

@@

We will use the change of variables $u(t,x)=e^{-bt}v(t,x)$

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
##Done: 1, 2, 4, 5, 6.

##Not Done: 3

@@

After setting $y=-s, dy=-ds$ in the integral of $\psi$:
$$u(-x,t)= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(-y)dy}=-u(x,t).$$
Therefore, using the uniqueness of the wave equation, odd initial conditions imply that the solution of the wave equation is odd.

@@

$$u(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy $$
Which means, when $x=-x$:

$$u(-x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(-x-y)^2/4kt}f(y)\;dy $$
Now use the change of variable $y=-s, dy=-ds$:

$$u(-x,t)= \frac{-1}{\sqrt{4 \pi kt}} \int_{\infty}^{-\infty} e^{-(-x+s)^2/4kt}f(-s)\;ds $$

@@

Now assume $f$ is odd. By definition, we have $f(-s)=-f(s)$, so

@@

And therefore our solution is odd when our initial $f(x)$ is an odd function. Similarly, if we assume $f(x)$ is even, by definition we have $f(-s)=f(s)$, so
$$u(-x,t)=\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds = 
\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(s)\;ds = u(x,t), $$
meaning (using the uniqueness of the heat equation), our solution is also even.

@@

\begin{equation}\label{eqn:tderiv}
u_t = -af'(x-at), \quad u_{tt}=a^2 f''(x-at)
\end{equation}

\begin{equation}\label{eqn:xderiv}
u_x = f'(x-at), \quad u_{xx}=f''(x-at)
\end{equation}

If $u(x, t) = f(x-at)$ is a solution to the wave equation, then we must have $u_{tt}=c^2 u_{xx}$. By substituting (\ref{eqn:tderiv}) and (\ref{eqn:xderiv}) into the PDE, we get

$$a^2 f''=c^2f'' ,$$
which implies at $a^2=c^2$, or $a= \pm c$. However, if $f$ is a linear function, then the second derivative $f''=0$ and we cannot conclude that $a^2=c^2$.

@@

If $u(x, t) = f(x-at)$ is a solution to the wave equation, then we must have $u_{t}=k u_{xx}$. By substituting (\ref{eqn:tderiv}) and (\ref{eqn:xderiv}) into the PDE, we get

$$-af'=kf'' $$
$$-af'=kf'' $$
$$ \frac{-a}{k}= \frac{f''}{f'} $$
Now if we set $\xi=x-at$, 
$$\frac{-a}{k}=\frac{f''}{f'}= \frac{d}{d\xi} \ln(f') $$
So integrating with respect to $\xi$,
$$\frac{-a}{k}\xi+C= \ln (f')$$
$$ f'(\xi) = e^{ \frac{-a}{k} \xi+C_1} = C_2 e^{\frac{-a}{k} \xi} $$
Now if we integrate again with respect to $\xi$,
$$f(\xi)=\frac{-k}{a}C_2 e^{\frac{-a}{k}\xi}+C_3 $$

Set a new constant $C_4=\frac{-k}{a}C_3$,
$$f(\xi)=C_4 e^{\frac{-a}{k}\xi}+C_3 $$

If we change back to the original variables,
$$f(x-at)=C_4 e^ \frac{-a(x-at)}{k}+C_3 $$

Thus, we found the solution
\begin{equation} \label{eqn:findf}
u(x, t) = f(x-at)=A e^{\frac{-a}{k}(x-at)}+B,
\end{equation}
where $A$ and $B$ are constants.

We can check that this satisfies the heat equation:
$$u_t=A\frac{a^2}{k} e^ {\frac{-a}{k}(x-at)}$$

$$u_x=A\frac{-a}{k} e^ {\frac{-a}{k}(x-at)}$$

$$u_{xx}=A\frac{a^2}{k^2} e^ {\frac{-a}{k}(x-at)}$$
$$u_t-ku_{xx}=A\frac{a^2}{k} e^ {\frac{-a}{k}(x-at)}-kA\frac{a^2}{k^2} e^ {\frac{-a}{k}(x-at)}=0$$
However, (\ref{eqn:findf}) will be a solution regardless of the value of $a$; i.e., $a$ is arbitrary.

## Question 3 (Heat Equation with Robin boundary)
Consider the following problem with a Robin boundary condition:
\begin{equation*} (\ast)
    \begin{cases}
      u_{t}=k u_{xx} & 0 <x<\infty,0 <t<\infty \\
      u(0,x)=x & t=0,0 <x<\infty \\
      u_x(0,t)-2u(0,t)=0 & x=0
    \end{cases}       
\end{equation*}

The purpose of this exercise is to verify the solution formula for $(\ast)$. Let $f(x)=x$ for $x>0$ and let $f(x)=x+1-e^{2x}$ for $x<0$, and let

$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy. $$
**(a)** What PDE and initial conditions does $v(x,t)$ satisfy for $-\infty < x < \infty$?

$v(x,t)$ satisfies

\begin{equation*}
    \begin{cases}
      u_{t}=k u_{xx} & 0 <x<\infty,0 <t<\infty \\
      u(0,x)=x & t=0,0 <x<\infty \\
    \end{cases}       
\end{equation*}

**(b)** Let $w=v_x-2v$. What PDE and initial condition does $w(x,y)$ satisfy for $-\infty < x < \infty$?

**(c)** Show that $f'(x)-2f(x)$ is an odd function for $x \neq 0$.
\begin{equation*}  f(x)=
\begin{cases}
   x & \text{ if } x >0 \\
   x+1-e^{2x} & \text{ if } x<0\\
\end{cases}
\end{equation*}

\begin{equation*}  f'(x)=
\begin{cases}
   1 & \text{ if } x >0 \\
   1-2e^{2x} & \text{ if } x<0\\
\end{cases}
\end{equation*}

\begin{equation*}  f'(x)-2f(x)=
\begin{cases}
   1-2x & \text{ if } x >0 \\
   -1-2x & \text{ if } x<0\\
\end{cases}
\end{equation*}

Which means that
\begin{equation*}  f'(-x)-2f(-x)=
\begin{cases}
   1-2(-x) & \text{ if } (-x) >0 \\
   -1-2(-x) & \text{ if } (-x) <0\\
\end{cases}
\end{equation*}

But this is equivalent to
\begin{equation*}
f'(-x)-2f(-x)=
\begin{cases}
   1+2x & \text{ if } x <0 \\
    -1+2x & \text{ if } x>0\\
\end{cases}
\end{equation*}

which is exactly $-(f'(x)-2f(x))$. Therefore, $f'(x)-2f(x)$ is an odd function of $x$.

**(d)** Use Exercise 2.4.11 to show that $w$ is an odd function of $x$.

**(e)** Deduce that $v(x, t)$ satisfies $(\ast)$ for $x>0$. Assuming uniqueness, deduce that the solution of $(\ast)$ is given by
$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy. $$

@@

Define $w(t,x)=u(t,x)-v(t,x)$ and note that $w$ solves the heat equation in $\{ 0 \leq x \leq l, 0 \leq t \leq T \}$:
$$w_t-kw_{xx}=u_t-v_t-k(u_{xx}-v_{xx})=(u_t-ku_{xx})-(v_t-kv_{xx})=0-0=0 $$
Since we are assuming $u \leq v$ for $t=0,x=0,$ and $x=l$, then $w \leq 0$ for $t=0,x=0,$ and $x=l$. But by the Maximum Principle, $w$ assumes its maximum value on one of those three lines. This means we can take $T$ to be arbitrarily large and still have $w \leq 0$, i.e., $w \leq 0$ on $\{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$ and therefore $u \leq v$ for $0 \leq t \leq \infty$ and $0 \leq x \leq l$.

@@

We will use the change of variables $u(t,x)=e^{-bt}v(t,x)$, and differentiate to find that:

$$u_t=e^{-bt}v_t-be^{-bt}v$$
$$u_x=e^{-bt}v_x$$
$$u_{xx}=e^{-bt}v_{xx}$$

So $u_t-ku_{xx}+bu=0$ becomes

$$e^{-bt}v_t-be^{-bt}v-k \left(e^{-bt}v_{xx}\right) + be^{-bt}v=0 $$
After canceling the $be^{-bt}v$ terms and dividing by $e^{-bt}$ (which we can do because it cannot equal zero), we are left with

$$v_t-kv_{xx}=0, \quad v(0,x)=u(0,x)=\psi(x) $$
Hence, the general solution of $v$ is

$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy, $$
and the general solution of $u$ is therefore

$$u(x,t)= \frac{e^{-bt}}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy. $$
